---
title: "Work Log: Registration of Paul Smith's Shorebird Tag Database with sensorgnome.org and motus-wts.org"
author: "John Brzustowski"
date: "December, 2015"
output:
   pdf_document :
       toc: true
       toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment="")
```
# Summary

Paul Smith has assembled a spreadsheet giving metadata for deployments
of coded ID tags on North American shorebirds from 2012 to 2015.  This
document records how this spreadsheet was reconciled with existing tag
registrations at sensorgnome.org and motus-wts.org.  As much as
possible, the steps here were coded as functions in the _motus_ R
package, so that other projects could more easily get their tag
metadata into motus-wts.org

# Preliminaries

First, read the list of projects from motus:

```{r ref.label="functions"}
```

```{r}
library(motus, quietly=TRUE)
motusLoadSecrets("~/.secrets/motusSecrets.json")

projs = motusListProjects()
```

Next, read the spreadsheet.  We copied the entire sheet and pasted as
values into a new sheet, which was saved as a .CSV file.  We had tried
to read the sheet using the `readxl::read_excel()` function, but this
did not work well because in some columns (e.g. TrackerID), values
were present as both numbers and as textual output of formulas.  The
current version of `read_excel()` function doesn't handle a
heterogenous column very well.

```{r}
x = read.csv("paul_smith_shorebird_tag_metadata_noformatting.csv", as.is=TRUE)
x = as.tbl(x)
x


```

Make some changes we'll need later:
```{r warning=FALSE}
x = x %>%
    mutate(
        ## fix burst intervals
        Burst = Burst %>%
            ## drop tildes: e.g.change ~6 to 6
            sub("^~", "", .) %>%
            ## change ranges to integer portion of 2nd value
            ## e.g. 5.9 / 6.1 -> 6
            sub("([0-9]+\\.[0-9]+) / (([0-9]+)\\.[0-9]+)", "\\3", perl=TRUE, .) %>%
            as.numeric
    )
             
```

Match names in the spreadsheet to motus projects, using a simple gui:

```{r warning=FALSE}

pmapFile = "projectMap.rds"
if (file.exists(pmapFile)) {
    pmap = readRDS(pmapFile)
} else {
    pmap = gmatch("Match Spreadsheet Names with Motus Projects",
                  "Spreadsheet Name",
                  unique(x$name),
                  "Motus Project",
                  paste(projs$id, projs$name))
    if (! is.null(pmap)) {
        ## extract projID
        names(pmap)=c("spreadsheetName", "motusProject")
        pmap$motusID = as.numeric(sub(" .*", "", pmap$r))
        saveRDS(pmap, "projectMap.rds")
    }
}
## inverse project map, from motusID to spreadsheetname
pmapi = character()
pmapi[pmap$motusID] = pmap$spreadsheetName
```

This leads to the following correspondence

```{r, echo=FALSE}
pmap = as.tbl(pmap)
pmap
## make a lookup table
projID = structure(as.list(pmap$motusID), names = pmap$spreadsheetName)
## add some entries explicitly
projID[["Kate Obrien"]] = 25
projID[["EBM Coats"]] = 66

```

Now get the full tag databases from sensorgnome.org for 2012-2015

```{r, cache=TRUE}
sg = rbind(
    cbind( year=2012, read.csv("2012_tags.csv", as.is=TRUE)),
    cbind( year=2013, read.csv("2013_tags.csv", as.is=TRUE)),
    ## drop two extra columns from recent tag DBs
    cbind( year=2014, read.csv("2014_tags.csv", as.is=TRUE)[, 1:15]),
    cbind( year=2015, read.csv("2015_tags.csv", as.is=TRUE)[, 1:15])
) %>% as.tbl

sg
```

# Tags Already Registered With Motus

A number of tags have already been registered with motus.  Which ones?
 
Work project by project.

```{r, cached=TRUE, warning=FALSE}
found = multi = collide = lost = NULL

## sets of motus-registered tags
motus = NULL

for (p in pmap$spreadsheetName) {
    ## lookup the motus projectID
    pid = projID[[p]]
    
    ## get tags registered for this project with motus
    motusp = getProjectTags(pid)

    ## look just at spreadsheet tags for this project
    xp = x %>% filter(name == p)

    ## match spreadsheet tags to motus tags
    m = matchTags(xp, motusp, "TrackerID", "Burst", "Yr")

    found = found %>% rbind( m$unique )
    multi = multi %>% rbind( m$multi )
    collide = collide %>% rbind (m$collide)
    lost  = lost  %>% rbind( m$none )
    
    ## add motus tags from this project to full set
    motus = motus %>% rbind (motusp)
}

```
## Results of Initial Matching

Out of `r nrow(x)` tags,

 * `r nrow(found) %or% 'none'` had unique matches in the motus registry
 * `r (nrow(multi) - sum(duplicated(multi[, 1:ncol(x)])))  %or% 'none'` had multiple matches; i.e. matched to more than one tag in the motus database.
 * `r nrow(collide) %or% 'none'` collided; i.e. matched to the same
 motus tag as at least one other user tag.  These will be saved in
 `2015_paul_smith_shorebird_tag_collisions.rds`. We need to determine
 which of these are typos, and which are true re-deployments of
 tags. **[Sjoerd]**
 * `r nrow(lost) %or% 'none'` had no matches; for these, the distribution across name and year is:

```{r}
table(lost$name, lost$Yr)

```
## Registrations to 'wrong' project

We look for registrations of tags to the wrong project.
First, remove the unique matches and collision from the motus set
```{r}
used = c(found$tagID, collide$tagID)
motus = motus %>% subset(! tagID %in% used)
```
Now, get additional registration from projects "EBM Coats" and "Obrien":

```{r}
motus = motus %>%
    bind_rows(getProjectTags(projID[["EBM Coats"]])) %>%
    bind_rows(getProjectTags(projID[["Kate Obrien"]]))
```

Next, look for 'lost' tags across all projects in the corresponding year.

```{r}
m2 = matchTags(lost, motus, "TrackerID", "Burst", "Yr")
remap = m2$unique %>% mutate(newProjectID = unlist(projID[name]))
lost = m2$none
```
This leads to a total of `r nrow(m2$unique)` tags which need to be reassigned
between projects, by year:
```{r}
table(remap$name,pmapi[remap$projectID], remap$Yr, dnn=c("New Name", "Original Name", "Year"))
## save the remap set, if not already generated on a previous run
reassignFile = "2015_paul_smith_shorebird_tag_reassign_projects.rds"
saveRDS(remap, reassignFile)
## write to small .csv file for Denis Lepage
remap %>%
    select(tagID, projectID, newProjectID) %>%
    write.csv("2015_paul_smith_shorebird_tag_reassign_project.csv", row.names = FALSE)
```

There are still `r nrow(lost)` tags not matched:

```{r}
table(lost$name, lost$Yr)
```

## Christian Friis Tags 2014-2015

These tags are not found:

```{r}
P = "Christian Friis"
pid = projID[[P]]

cf = lost %>% filter( name==P) %>% arrange(Yr, TrackerID)
cf %>% select(Yr, TrackerID, Burst)

```

#### Tag 172
This tag was missing on the first run, but has since been registered.
Here is the code used:

```{r, cached=TRUE}
## only run once

if (172 %in% as.numeric(cf$mfgID)) {
    ## Tag 172 from 2015 was a twin of that from project Selva, so dropped.
    ## Read its parameters from the database of James Bay project tags
    tmp = read.csv("/disco/SG/contrib/2015/JamesBay/2015_JamesBay_tag_database.csv", as.is=TRUE) %>%
        subset(id == 172)

    regts = ymd("2015-05-30")

    with(tmp, 
         motusRegisterTag(
             projectID = pid,
             mfgID = "172",
             manufacturer = "Lotek",
             type = "ID",
             codeSet = "Lotek-4",
             offsetFreq = dfreq + (fcdFreq - tagFreq) * 1000,
             period = bi,
             periodSD = bi.sd,
             pulseLen = 2.5,
             param1 = g1,
             param2 = g2,
             param3 = g3,
             param4 = g1.sd,
             param5 = g2.sd,
             param6 = g3.sd,
             paramType = 1,
             ts = as.numeric(regts),
             nomFreq = tagFreq,
             dateBin = sprintf("%4d-%1d", year(regts), ceiling(month(regts)/3)),
             show = TRUE
         ))
}

```

### Tags 376-379
These were registered, but the burst interval is ~9.7 s, not ~5.9s
as reported in the spreadsheet. **[Sjoerd]**

```{r, cached=TRUE}
bindUserMotus = function(u, m) {
    ## bind rows of user tags to motus tags, respectively
    if (nrow(u) != nrow(m))
        stop("need same number of rows in both parameters")
    ## lookup specified user rows in table of missing rows
    src = which(duplicated(rbind(u, lost))) - nrow(u)
    if (length(src) != nrow(u) | any(src < 1 | src > nrow(lost)))
        stop("specified user rows do not come from table 'lost'")
    r = cbind(u, m)
    names(r)[names(r)=="comment"] = c("comment.x", "comment.y")
    found <<- found %>% bind_rows(r)
    lost <<- lost[- src,]
    motus <<- motus %>% subset( ! tagID %in% m$tagID)
}

biFixId = 376:379


bindUserMotus(
    cf %>% filter(as.integer(TrackerID) %in% biFixId) %>% arrange(TrackerID),
    subset(motus,
           projectID==projID["Christian Friis"] &
           as.integer(mfgID) %in% (376:379)
           ) %>% arrange(mfgID)
)

```

The remaining 3 Friis tags, #226, 301, and 338 from 2014 with BI ~6s
have not been registered with sg or motus:

```{r}
needReg = subset(cf, as.integer(TrackerID) %in% c(226, 301, 338))
needReg %>% select(name, Yr, Month, Day, LocationID, TrackerID, Burst)
```

The deployment location matches the Piskwamish site.  226 shows up
with a BI of 5.8983s there, but neither 301 nor 338 is present.
Perhaps these two had mis-recorded ID numbers?  We register tag 226:

```{r, cached=TRUE}
    ## ## Read its parameters from the database of James Bay project tags
    ## tmp = read.csv("/disco/SG/contrib/2015/JamesBay/2015_JamesBay_tag_database.csv", as.is=TRUE) %>%
    ##     subset(id == 172)

    ## regts = ymd("2015-05-30")

    ## with(tmp, 
    ##      motusRegisterTag(
    ##          projectID = pid,
    ##          mfgID = "172",
    ##          manufacturer = "Lotek",
    ##          type = "ID",
    ##          codeSet = "Lotek-4",
    ##          offsetFreq = dfreq + (fcdFreq - tagFreq) * 1000,
    ##          period = bi,
    ##          periodSD = bi.sd,
    ##          pulseLen = 2.5,
    ##          param1 = g1,
    ##          param2 = g2,
    ##          param3 = g3,
    ##          param4 = g1.sd,
    ##          param5 = g2.sd,
    ##          param6 = g3.sd,
    ##          paramType = 1,
    ##          ts = as.numeric(regts),
    ##          nomFreq = tagFreq,
    ##          dateBin = sprintf("%4d-%1d", year(regts), ceiling(month(regts)/3)),
    ##          show = TRUE
    ##      ))
##}

```

## Larry Niles Tags 2014 - 2015

These tags are not found:

```{r}
P = "Larry Niles"
pid = projID[[P]]
t = lost %>% filter( name==P) %>% arrange(Yr, TrackerID)
t %>% select(Yr, TrackerID, Burst)

```
But tags 244...261 and 511 are registered to Niles 2015 with BI ~6.1, and these
are otherwise unmatched, so the BI in the spreadsheet must be incorrect.
**[Sjoerd]**

```{r}

fixBI = c(244, 246:261, 511)

bindUserMotus(
    t %>% filter(as.integer(TrackerID) %in% fixBI),
    subset(motus, projectID == projID["Larry Niles"] &
                  as.integer(mfgID) %in% fixBI
           ) %>% arrange(mfgID)
    )

```
These are still these unmatched:

```{r}
t = lost %>% filter( name==P) %>% arrange(Yr, TrackerID)
t %>% select(Yr, TrackerID, Burst) %>% as.data.frame
```

There are still 2 unmatched Niles tags; these have unknown IDs:
```{r}
subset(t, ! as.integer(TrackerID) %in% fixBI)
```
These might be among the `r sum(sg$proj=="Niles") - sum(found$name=="Larry Niles")`
tags registered with sensorgnome to the Larry Niles project but for which
no metadata were provided in Paul's spreadsheet.

## Paul Smith 2015 Tags

There is a single tag left with __invalid__ ID 545:

```{r}
lost %>% filter(name=="Paul Smith") %>% as.data.frame
```

Perhaps it is either 454 or 445, both registered to motus project
_EBM Coats_, and not matched to any other tags in the spreadsheet.
**[Sjoerd]**


## Rebecca Holberton / Kate Obrien / Lindsay Tudor 2015, 2014 Tags

These tags are still unmatched:

```{r}
P = "Lindsay Tudor"
pid = projID[[P]]
t = lost %>% filter(name==P & Yr %in% c("2014", "2015")) %>%
    arrange(Yr, TrackerID)
t %>% select(Yr, TrackerID, Burst)
```
Tags 274, 275, and 276 apparently have the wrong Burst interval in the
spreadsheet, which lists 10.2 instead of the 8.9 common to the rest of
the tags in that set in the spreadsheet, and in Kate Obrien's registrations.

We match these:
```{r}

fixBI = 274:276

bindUserMotus(
    t %>% filter(as.integer(TrackerID) %in% fixBI),
    subset(motus, projectID == projID["Kate Obrien"] &
                  as.integer(mfgID) %in% fixBI
           ) %>% arrange(mfgID)
    )

```
This leaves only a single tag for _Lindsay Tudor_ in 2014 or 2015, namely
this one with no ID:

```{r}
lost %>% filter(name=="Lindsay Tudor" & Yr %in% c(2014,2015)) %>%
    select(name, Yr, LocationID, TrackerID, Burst)
```
It should be possible to find this one with a full search of POPH site
data.

## Leftover Tags

This leaves `r nrow(lost)` tags still not matched:

```{r}
table(lost$name, lost$Yr)
```

## Tag Collisions

Some tags in the spreadsheet appear indistinguishable, and it will take some
digging to determine whether this is really the case, or whether it is
due to typos:  **[Sjoerd]**

```{r}
collide = collide %>% arrange (name, Yr, TrackerID)
collide %>% select(name, Yr, Month, Day, TrackerID, Burst)
     %>% as.data.frame

saveRDS(collide, "2015_paul_smith_shorebird_tag_collisions.rds")
```

### Yves Aubry Tag Collisions
For the two Aubry tags, it appears one of them should
be listed as #203, as this ID was registered by him for 2015, but
does not appear in the spreadsheet.

### Larry Niles Tag Collisions

#### ID = 314
For tags 314 and 314.1, the burst intervals should be different.
There are two tags with ID 314 registered with sensorgnome by Larry Niles
in 2015; one had BI 6.1s, and we refered to it as 314.1; the other had BI 5.3s
and we refered to it as 314.0.  It appears the spreadsheet has the incorrect
BI for 314 (i.e. for 314.0), as the deployment date and location match detections
for the tag 314 with BI=5.3s

#### ID = 345

There is a tag with BI=5.3 sec and ID=345 registered to Larry Niles
in 2015, and detected at a number of his sites.  I'm not sure what
the second tag 345 really is.

### Christian Friis Tag Collisions

For the two Friis tags #183

```
\newpage

```{r, child=c("appendix1.Rmd","appendix2.Rmd","appendix3.Rmd", "functions.Rmd")}
```

