---
title: "Lotek Tag Lifespan Model"
author: "John Brzustowski"
date: "4 April, 2016"
output:
   pdf_document :
       toc: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(comment="")
options(warn=0)
library(dplyr)
library(lattice)
```

We need some reasonable upper bound for the lifetimes of tags,
by battery and BI.

Lotek provides a short table here:

   http://www.lotek.com/bird+bat-nano.pdf  (As of 1 April, 2016)

and there are additional models listed in "ANTC Spec Sheet.pdf" (no
source), including the only numbers for battery type "3-1".  Since
lifetime is listed as depending only on the battery (numeric portion
of model string), we model on that.

```{r}
lt = as.tbl(read.csv("./lotekTagLifespanByBatteryAndBI.csv"))
```

Add inverses of columns bi and dutyCycle, for modelling

```{r}

ls = lt %>% mutate(biInv = 1.0 / bi, dcInv = 1.0 / dutyCycle)

print(ls)
```

A simple model assumes that the battery capacity, $K$, depends on the
battery model, and that tag power consumption $r$ is a sum of a baseline
rate $r_0$ plus a pulse-dependent rate $r_1$.  All tags transmit 4 pulses per burst,
so the pulse-dependent rate $r_1$ depends only on duty cycle and burst interval
like so:  $r_1 = r_p * \frac{dutyCyle}{BI}$

The full non-linear model is :

\[
lifespan = \frac{K}{r_0 + \frac{r_p * dutyCycle}{BI}}
\]

This is over-parameterized (we can divide top and bottom by $r_0$ to
get a model with two parameters), so we simplify by rewriting $K/r_0$
as $D$, the number of days of battery life at the baseline rate, and
$r_p / r_0$ be $r_t$, the relative rate of power consumption during
transmission, versus baseline.  The new model is:

\[
lifespan = \frac{D}{1 + \frac{r_t * dutyCycle}{BI}}
\]

We fit the model to each type of battery:

```{r, show=TRUE, results=FALSE, output=FALSE}
par = res = pred = NULL
bnames = unique(ls$battery)
for (m in bnames) {
    res[[m]] = nls(lifespan~D / (1 + rt * dutyCycle / bi),
                   subset(ls, battery==m), list(D = 500, rt = 5))
    par = rbind(par, c(coefficients(res[[m]]), max(residuals(res[[m]]))))
    pred = rbind(pred,
                 data.frame(
                     battery=m,
                     bi=1:40,
                     lifespan=predict(res[[m]], list(bi=1:40, dutyCycle=1))
                 ))
}
rownames(par) = bnames
colnames(par) = c("D", "rt", "Max Residual (days)")
```

Stu Mackenzie pointed out there is a light-weight version of the *NTQB-1*
which he's dubbed *NTQB-1-LW*, with "~2/3 the lifetime of the NTQB-1".
For now we'll assume that means the $D$ parameter for that model is 2/3
that for the NTQB-1

```{r}
par = rbind(c(par[["1","D"]] * 2 / 3, par[["1", "rt"]], NA), par)
rownames(par)[1] = "1-LW"
pred = rbind(data.frame(
    battery="1-LW",
    bi=1:40,
    lifespan=par[["1-LW", "D"]] / (1 + par[["1-LW", "rt"]] / (1:40)))
  , pred)
```

Now map model names to batteries.  We do this as a simple table
because the naming scheme isn't consistent enough to bother doing it
programmatically.

```{r}
## Models and the batteries they correspond to

modelBattery = list(
"NTQB-1-LW"  = "1-LW",
"NTQB-1"     = "1",
"NTQB-2"     = "2",
"NTQB-3-2"   = "3-2",
"NTQB-4-2"   = "4-2",
"NTQB-6-1"   = "6-1",
"NTQB-6-2"   = "6-2",
"NTQBW-3-2"  = "3-2",
"NTQBW-2"    = "2",
"NTQBW-4-2"  = "4-2",
"NTQBW-6-2"  = "6-2",
"ANTC-M1-1"  = "1",
"ANTC-M2-1"  = "2",
"ANTC-M3-1"  = "3-1",
"ANTC-M3-2"  = "3-2",
"ANTC-M4-2"  = "4-2",
"ANTC-M4-2S" = "4-2",
"ANTC-M4-2L" = "4-2",
"ANTC-M6-1"  = "6-1",
"ANTC-M6-2"  = "6-2",
"ANTCW-M1-1"  = "1",
"ANTCW-M2-1"  = "2",
"ANTCW-M3-1"  = "3-1",
"ANTCW-M3-2"  = "3-2",
"ANTCW-M4-2"  = "4-2",
"ANTCW-M4-2S" = "4-2",
"ANTCW-M4-2L" = "4-2",
"ANTCW-M6-1"  = "6-1",
"ANTCW-M6-2"  = "6-2",
"ACT-521" = "521",
"ACT-626" = "626",
"ACT-393" = "393"
)

modPar = NULL
for (b in names(modelBattery))
    modPar = rbind(modPar, par[modelBattery[[b]],])
rownames(modPar) = names(modelBattery)
```

The results show good agreement with the data table from Lotek:

```{r}
print(round(par, 1))
```


The parameter $r_t$ can be interpreted as the ratio of energy consumed
during 1 second with a burst to that consumed during 1 second without
a burst.  Estimates of this parameter vary only by 5% across tag
types, and in monotonic fashion, perhaps due to variation in battery
internal resistance.  The table provided by Lotek only covers $2 <= BI
<= 20$ (the larger value from BI=10s @ 50% duty cycle); curves are extrapolated
down to 1s and from 20 to 40s using the fitted model.

```{r}
xyplot(log10(lifespan)~bi, groups=battery, pred,
       auto.key=list(corner=c(.6,.1), columns=3),
       main="Reported (X) and Predicted (o) Tag Lifespan by Battery Type",
       xlab="Burst Interval (seconds)",
       ylab="Lifespan (log10(days); 1->10, 2->100, 3->1000)",
       type="b",
       panel = function(x, y, type, groups, ...) {
           panel.xyplot(x, y, type, groups, ...)
           meas = which(x %in% c(2, 5, 10, 20) & groups != "1-LW")
           panel.points(x[meas], y[meas], pch="X", cex=1.5)
           panel.abline(h=0, lty=2, col="gray")
           }
       )
```

```{r, include=FALSE}
tagLifespanPars = modPar[, 1:2]
save(tagLifespanPars, file="modelLotekTagLifeSpan.rda")
